---
title: "Tidy Tuesday"
---

***

<a href="https://github.com/rfordatascience/tidytuesday" target="_blank">Tidy Tuesday</a> is a weekly data project aimed at the R users. It is organised by the <a href="https://twitter.com/R4DScommunity" target="_blank">R4DS</a> community.  
It is a great way of practicing with differnt types of data and using different techniques. Wrangling, tidying, plotting and modelling different types of data.  

One my latest attempt was in
<a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-09" target="_blank">week 28 (09/10/2018)</a>,
which was voter turnout data from 1980 to 2014 for the USA. It included the number of age-eligible voters versus total ballots/votes counted by state and year. I decided to try and use the
<a href="https://github.com/thomasp85/gganimate" target="_blank">gganimate</a>
library to create an animated plot.

<br>  

### Previous Weeks
[Week 30 - Movie Data](tidytuesday/week30.html)  
[Week 28 - US Voter Data](tidytuesday/week28.html)  
[Week 5 - US Census Data](tidytuesday/week5.html)

<br>  

### Week 28  

The animation shows the number of elgible voters versus the number of votes over time.  

```{r week28_code, warning=FALSE, message=FALSE, error=FALSE, fig.align='center',echo=FALSE}
library(tidyverse)
library(gganimate)
library(scales)

 # Load data
voter_data <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-10-09/voter_turnout.csv")

 # Remove unwanted entries
state_data <- voter_data %>% 
  filter(!state %in% c("United States", "United States (Excl. Louisiana)"))

 # Create base plot for animation
votes_vs_eligible <- ggplot(state_data, aes(votes, eligible_voters, colour = state)) +
  geom_point(show.legend = FALSE) + theme_light() +
  scale_y_continuous(labels = comma_format()) +
  scale_x_continuous(labels = comma_format()) +
  labs(title = 'Eligible voters vs Number of votes', 
       subtitle = 'Year: {frame_time}',
       y = 'Eligible voters', x = 'Votes') +
  # extra lines needed for animation are below
  transition_time(year) +
  ease_aes('quartic-in-out')  # determines how points will move

animate(votes_vs_eligible, fps = 7.5)  # create the animation
```



### Code
```{r ref.label="week28_code", eval=FALSE}
```
