---
title: "2019 - Week 46 - CRAN"
params:
  top_n: 
    label: "Top n languages"
    value: 20
    input: slider
    min: 1
    max: 100
    step: 1
  perc_comparison:
    label: "Language to Compare"
    value: R
    input: select
    choices: ["R", "C", "HTML", "C++", "SVG", "JavaScript"] 
---


***

A copy of the RMD file can be dowloaded from [here](www.aboland.ie)

***

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(dplyr)
cran_code <- 
  readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-12/loc_cran_packages.csv")
```

**Note**: `params$top_n` and `params$perc_comparison`, which are used throughout are markdown parameters, for this specific rendering they are `r params$top_n` and "`r params$perc_comparison`" respectively.  
See the [RMarkdown with Parameters] sections at the end for more info.

## Top `r params$top_n` Laguages

```{r top n by package, warning=FALSE, fig.align='center'}
cran_code %>%
  group_by(language) %>%  # group by langauge
  summarise(lang_count = n(),  # calc number of packages by language
            lang_lines = sum(code)  # calc number of lines of code by language
            ) -> cran_language_summary

cran_language_summary %>%
  top_n(params$top_n, lang_count) %>%  # filter by top n using params$top_n
  ggplot(aes(x = reorder(language, lang_count, decreasing = T) , y = lang_count)) +
  geom_histogram(stat = "identity", fill = "steelblue") +
  scale_y_continuous(labels = comma) +  # display large numbers with a comma
  coord_flip() +  # flip plot to see language names better
  labs(title = paste0("Top ", params$top_n," Languages on CRAN (by number of packages)"), y = "Number of Packages", x = "")
```



```{r top-n-by-lines-code, warning=FALSE, fig.align='center'}
cran_language_summary %>%
  top_n(params$top_n, lang_lines) %>%  # filter by top n using params$top_n
  ggplot(aes(x = reorder(language, lang_lines, decreasing = T) , y = lang_lines)) +
  geom_histogram(stat = "identity", fill = "steelblue") +
  scale_y_continuous(labels = comma) +
  coord_flip() +
  labs(title = paste0("Top ", params$top_n," Languages on CRAN (by lines of code)"), 
       x = "", y = "Lines of Code")
```


## % of `r params$perc_comparison` Used

```{r perc-r-in-use, warning=FALSE, echo=FALSE, include=FALSE}
# must be a neater way of calculating this
cran_code %>%
  group_by(pkg_name) %>%  # group by package name
  summarise(total_lines = sum(code)) %>%  # calc total lines
  left_join(
    cran_code %>%
      filter(language == params$perc_comparison) %>%  # filter R only
      select(pkg_name, code),  # select relevant variables
    by = "pkg_name"  # join using `pkg_name`
    ) %>%
  mutate(code = ifelse(is.na(code), 0, code),  # change na's to 0 
         perc_lang = code/total_lines) %>%  # calc % of R
  na.omit() -> cran_perc_lang

cran_perc_lang %>%
  ggplot(aes(perc_lang)) +
  geom_histogram(fill = "steelblue", colour = "black", binwidth = 0.1) +
  scale_x_continuous(labels = percent) +  # use % on x axis
  labs(title = paste0("% of ", params$perc_comparison, " used in packges on CRAN (", nrow(cran_perc_lang), " packages)"), 
       x = "", y = "No. of Packages")
```


Out of the `r cran_code$pkg_name %>% unique() %>% length() %>% prettyNum(big.mark = ",")` packages,  `r length(unique(cran_code$pkg_name)) - nrow(cran_perc_lang)` package has no code, which leaves `r nrow(cran_perc_lang) %>% prettyNum(big.mark = ",")` packages which contain some code. Of those packages containing some code `r sum(cran_perc_lang$perc_lang == 0)` contain no R code and `r sum(cran_perc_lang$perc_lang == 1) %>% prettyNum(big.mark = ",")` packages contain only R code.


```{r perc-r-in-use, fig.align='center'}
```





<br><br><br><br>

## RMarkdown with Parameters

This document makes use of parameters. These can be included using the params argument in the preamble of the markdown document.

```{r, eval=FALSE}
---
title: "2019 - Week 46 - CRAN"
params:
  top_n:
    label: "Top n languages"
    value: 20
    input: slider
    min: 1
    max: 100
    step: 1
  perc_comparison:
    label: "Language to Compare"
    value: R
    input: select
    choices: ["R", "C", "HTML", "C++", "SVG", "JavaScript"]
---
```


`params$top_n` and `params$perc_comparison` can be used throughout the document as a variables.

The document can be knitted with the defalut values by clicking the `Knit` button or specific values using `rmarkdown::render("Week46_crand.Rmd, params = list(top_n = 20, perc_comparison = "R"))"`.

A user interface can be used by selecting `Knit with Parameters` from the dropdown menu for `Knit`, or `rmarkdown::render("MyDocument.Rmd", params = "ask")`.

See [here](https://bookdown.org/yihui/rmarkdown/params-knit.html) for more information on parameters.

