---
title: "Week 28 - 2018"
---

***

## Voter Data

<a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-10-09" target="_blank">Week 28</a>
of tidy Tuesday is a dataset containing the number of age-eligible voters versus total ballots/votes counted by state and year.

### Animation

Using the
<a href="https://github.com/thomasp85/gganimate" target="_blank">gganimate</a>
library, I created an animation over time of the number of eligible voters versus the number of votes.

```{r week28_code, warning=FALSE, message=FALSE, error=FALSE, fig.align='center', echo=FALSE, cache=TRUE}
library(tidyverse)
library(gganimate)
library(scales)

 # Load data
voter_data <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-10-09/voter_turnout.csv")

 # Remove unwanted entries
state_data <- voter_data %>% 
  filter(!state %in% c("United States", "United States (Excl. Louisiana)")) %>% 
  mutate(year = as.integer(year))

 # Create base plot for animation
votes_vs_eligible <- ggplot(state_data, aes(votes, eligible_voters, colour = state)) +
  geom_point(show.legend = FALSE) + theme_light() +
  scale_y_continuous(labels = comma_format()) +
  scale_x_continuous(labels = comma_format()) +
  labs(title = 'Eligible voters vs Number of votes', 
       subtitle = 'Year: {frame_time}',
       y = 'Eligible voters', x = 'Votes') +
  # extra lines needed for animation are below
  transition_time(year) +
  ease_aes('quartic-in-out')  # determines how points will move

animate(votes_vs_eligible, fps = 7.5)  # create the animation
```


## Code

```{r ref.label="week28_code", eval=FALSE}
```



<br>  
<br>  

### Twitter

<div align="center">
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Decided to try out gganimate for <a href="https://twitter.com/hashtag/TidyTuesday?src=hash&amp;ref_src=twsrc%5Etfw">#TidyTuesday</a> week 28. My gif shows eligible voters constantly increasing while the number of votes increases for a presidential election year. <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://twitter.com/hashtag/r4ds?src=hash&amp;ref_src=twsrc%5Etfw">#r4ds</a> <a href="https://t.co/mld76Dl4N7">pic.twitter.com/mld76Dl4N7</a></p>&mdash; Aidan Boland (@AidoBo) <a href="https://twitter.com/AidoBo/status/1050014566795612160?ref_src=twsrc%5Etfw">October 10, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
</div>
<br>
