---
title: "Week 32 - 2018"
---

***
  
## US Wind Turbine Data
<br>

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 9, fig.height = 5, 
                      fig.align = 'center',
                      warning = FALSE, error = FALSE, message = FALSE, cache = TRUE)
```

Tidy Tuesday <a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2017/2018-11-06" target="_blank">Week 32</a>
is wind turbine location and characteristic data from across the USA.  
A data dictionary can be found
<a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-11-06" target="_blank">here.</a>  
<br>  


## Code Walkthrough

For this weeks dataset, I decided to document my data exploration in an
<a href="https://rmarkdown.rstudio.com/" target="_blank">R markdown</a>
document.

#### Interesting sections
- [Interactive plots (plotly)](#InteractivePlots)  
- [Animations (gganimate)](#Animation)

<br>  


### Exploring the data

```{r libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(scales)
library(xtable)
theme_set(theme_light())  # Set ggplot theme globally
```
Having viewed the data dictionary, we see that missing values take three different forms, "n/a", "-9999", and "missing". These can be used when reading in the data.  
```{r read in data, warning=FALSE, message=FALSE}
raw_data <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-06/us_wind.csv", na = c("n/a", "-9999", "missing"))
```
<br>  

### Project Characteristics

The data contains `r nrow(raw_data)` data points and `r length(names(raw_data))` variables. There are `r length(unique(raw_data$p_name))` different projects in `r length(unique(raw_data$t_state))` states.  


Each project has characteristics, such as total capacity (MW), number of turbines and the year the project became operational.  

An intuitive relationship is the number of turbines and the capacity of a project. 

```{r project data plot1}
project_data <- raw_data %>% 
  group_by(p_name, p_year, p_tnum, p_cap, t_state) %>% 
  summarise(avg_cap = mean(t_cap, na.rm = T), 
            avg_height = mean(t_hh, na.rm = T), 
            avg_diameter = mean(t_rd, na.rm = T)) %>%
  ungroup()

project_data %>% 
  ggplot(aes(p_tnum, p_cap, colour = t_state)) + 
  geom_point() + 
  guides(colour = FALSE) +  # too many states to display in a legend
  scale_y_continuous(labels = number_format(suffix = " MW")) +
  labs(x = "number of turbines", y = "project capacity", title = "Capacity vs No. Turbines")
```
<br>  

The number of turbines can be used to determine the size of the points in a graph. This is a good way to include extra information in a graph. The project capacity, year, number of turbines and state are all included in the graph below, albeit the legend of which state is which colour is not displayed.
```{r project data plot2}
project_data %>% 
  ggplot(aes(p_year, p_cap, colour = t_state, size = p_tnum)) + 
  geom_point() + 
  guides(colour = FALSE, size = guide_legend("# turbines", reverse = T)) +
  scale_y_continuous(labels = number_format(suffix = " MW")) +
  labs(x = "year", y = "project capacity", title = "Capacity across time")
```

```{r project data plot3}
year_data <- raw_data %>% 
  group_by(p_year) %>% 
  summarise(avg_cap = mean(t_cap, na.rm = T), 
            avg_height = mean(t_hh, na.rm = T), 
            avg_diameter = mean(t_rd, na.rm = T), 
            avg_tnum = mean(p_tnum, na.rm = T)) %>%
  ungroup()

year_data %>% 
  ggplot(aes(p_year, avg_cap, size = avg_tnum)) + 
  geom_point() + 
  guides(size = guide_legend("Avg # turbines", reverse = T)) +
  scale_y_continuous(labels = number_format(suffix = " MW")) +
  labs(x = "year", y = "average project capacity", title = "Capacity across time")
```

<br>  

### Modelling project capacity

#### Collinearity
```{r}
 # model 1
project_lm <- lm(p_cap ~ p_tnum + avg_height + avg_diameter + p_year, data = project_data)
project_lm$coefficients
```
The height and diameter of turbines are going to be highly correlated, this is a classic example of collinearity. The Pearson correlation between average height and average diameter is `r round(cor(project_data %>% select(avg_height, avg_diameter), use = "pair")[1,2], 4)`.

#### Year as a continuous variable?

Height is removed, we now concentrate on the year variable. This can be thought of as a continuous variable or as a factor. 

```{r}
 # model 2 (year continous)
project_lm2 <- lm(p_cap ~ p_tnum + avg_diameter + p_year, data = project_data)
project_lm2$coefficients
```


```{r}
 # model 3 (year factor)
project_lm3 <- lm(p_cap ~ p_tnum + avg_diameter + as.factor(p_year), data = project_data)

# predictions
avg_diameter <- project_data %>% 
  filter(p_year == 2018) %>% 
  select(avg_diameter) %>% 
  colMeans(na.rm = T)

new_data_predicted <- data_frame(model = "Year as continuous", p_tnum = 1:140, avg_diameter, p_year = 2018) %>% 
  mutate(predicted_cap = predict(project_lm2, newdata = .)) %>% 
  bind_rows(
    data_frame(model = "Year as factor", p_tnum = 1:140, avg_diameter, p_year = 2018) %>%
      mutate(predicted_cap = predict(project_lm3, newdata = .))
    )

ggplot() +
  geom_point(aes(x = p_tnum, y = p_cap), data = project_data %>% filter(p_year > 2017, !is.na(p_cap))) +
  geom_line(aes(x = p_tnum, y = predicted_cap, colour = model), data = new_data_predicted) +
  labs(x = "number of turbines", y = "total capacity", 
       title = "Linear model of capacity", subtitle = "2018 prediction") + 
  scale_y_continuous(label = number_format(suffix = " MW"))
```

<br>  

It's worth fitting the model with only modern data. Using data from the year 2000 on fits very similar models for both continuous and factor models. In this case it's better to use year as a continuous variable because it reduces the number of parameters in the model.  
<br>  
```{r, echo = FALSE}
project_data_modern <- project_data %>% filter(p_year > 2000)
project_lm2_2 <- lm(p_cap ~ p_tnum + avg_diameter + p_year, data = project_data_modern)
project_lm3_2 <- lm(p_cap ~ p_tnum + avg_diameter + as.factor(p_year), data = project_data_modern)

new_data_lm2_2 <- data_frame(model = "Year as continuous", p_tnum = 1:140, avg_diameter, p_year = 2018) %>% 
  mutate(predicted_cap = predict(project_lm2_2, newdata = .))

new_data_combined2 <- new_data_lm2_2 %>% 
  bind_rows(
    data_frame(model = "Year as factor", p_tnum = 1:140, avg_diameter, p_year = 2018) %>% 
      mutate(predicted_cap = predict(project_lm3_2, newdata = .))
            )

ggplot() +
  geom_point(aes(x = p_tnum, y = p_cap), data = project_data %>% filter(p_year > 2017, !is.na(p_cap))) +
  geom_line(aes(x = p_tnum, y = predicted_cap, colour = model), data = new_data_combined2) +
  labs(x = "number of turbines", y = "total capacity", title = "Linear model of capacity", subtitle = "2018 prediction", colour = "model (>2010)") + 
  scale_y_continuous(label = number_format(suffix = " MW"))
```


```{r, results='asis'}
data_frame(Model = c("Year (cont)", "Year (fact)", "Modern Year (cont)", "Modern Year (fact)"),
           Rsq = c(
             cor(project_lm2$model$p_cap, project_lm2$fitted.values)^2,
             cor(project_lm3$model$p_cap, project_lm3$fitted.values)^2,
             cor(project_lm2_2$model$p_cap, project_lm2_2$fitted.values)^2,
             cor(project_lm3_2$model$p_cap, project_lm3_2$fitted.values)^2),
           RMSE = c(
             sqrt(mean(project_lm2$residuals^2)),
             sqrt(mean(project_lm3$residuals^2)),
             sqrt(mean(project_lm2_2$residuals^2)),
             sqrt(mean(project_lm3_2$residuals^2))
           )
) %>% xtable(digits = 4) %>% 
  print(type = "html", include.rownames = FALSE, html.table.attributes = list('align="center", border=1, width=40%'))
```

```{r, echo=F, eval=F}
summary(project_lm2)
summary(project_lm3)
summary(project_lm2_2)
summary(project_lm3_2)
```

<br>
<br>

### Turbine Characteristics

The data also contains information on different turbines, manufacturer, model, etc. The graphs below show the height and diameter versus capacity. The points are coloured by year, and done in 2 different ways to illustrate how a variable like year can be displayed as discrete or continuous.

```{r turbine data plots, warning=FALSE, fig.align='center'}
turbine_data <- raw_data %>% 
  group_by(t_manu, t_model, t_cap, t_hh, t_rd) %>% 
  summarise(year = min(p_year, na.rm = T)) %>%
  filter(!is.infinite(year)) %>% 
  ungroup()

turbine_data %>% ggplot(aes(t_cap, t_hh, colour = as.factor(year))) +
  geom_point() +
  scale_x_continuous(labels = number_format(suffix = " kW", big.mark = "")) +
  scale_y_continuous(labels = number_format(suffix = " m")) +
  labs(x = "turbine capacity", y = "turbine hub height", title = "Capacity vs Height", colour = "Year")

turbine_data %>% ggplot(aes(t_cap, t_rd, colour = year)) +
  geom_point() +
  scale_x_continuous(labels = number_format(suffix = " kW", big.mark = "")) +
  scale_y_continuous(labels = number_format(suffix = " m")) +
  labs(x = "turbine capacity", y = "turbine rotor diameter", title = "Capacity vs Rotor Diameter", colour = "Year")
```

#### Average height and diameter over time
```{r}
year_data %>% 
  gather(variable, measure, -p_year) %>% 
  filter(variable %in% c("avg_height", "avg_diameter")) %>% 
  ggplot(aes(p_year, measure, colour = variable)) + 
  geom_line() + 
  scale_y_continuous(labels = number_format(suffix = "m")) + 
  labs(x = "year", y = "measure", title = "Average Height and diameter of turbines over time", colour = NULL)
```



### Interactive plots {#InteractivePlots}

The
<a href="https://plot.ly/r/" target="_blank">plotly</a>
package is a great way to create interactive graphs. The `ggplotly` function will create an interactive version of a ggplot.


```{r project data plotly}
library(plotly)

plotly_1 <- project_data %>% 
  ggplot(aes(avg_cap, avg_height, colour = t_state, size = p_tnum, 
             text = paste0("Project: ", p_name, 
                           "<br>Year: ", p_year, 
                           "<br>State: ", t_state, 
                           "<br>Project Capacity: ", p_cap," MW", 
                           "<br>No. of Turbines: ", p_tnum))) +
  geom_point() +
  scale_x_continuous(labels = number_format(suffix = " kW")) + 
  scale_y_continuous(labels = number_format(suffix = " m")) +
  labs(x = "average turbine capacity", y = "average turbine height", 
       title = "Project's Turbine Capacity vs Height",
       colour = "State", size = NULL)
```
Double click on any state in the legend to display only the data from that state.
```{r project data plotly output}
ggplotly(plotly_1, tooltip = "text")
```
<br>

A 3d interactive plot can be useful, the user can explore the data themselves.  
```{r plotly 3d}

turbine_3d <- plot_ly(turbine_data %>% filter(year > 2010),  # too many observations without filter
                      x = ~t_hh, y = ~t_rd, z = ~t_cap, color = ~year,
                      type = "scatter3d", size = 0.5, hoverinfo = 'text',
                      text = ~paste("</br> capacity:", t_cap,  # this controls the information displayed when 
                                    "</br> height:", t_hh,     # hoovering over an observation
                                    "</br> diameter:", t_rd,
                                    "</br> year:", year)) %>% 
  layout(title = "Turbine heights and diameters versus capacity",
         scene = list(camera = list(eye = list(x = 1.8,y = 0.3,z = 0.28),
                                  center = list(x = 0, y = 0, z = -0.2)),  # set the default camera view
                    xaxis = list(title = "height", ticksuffix = " m"),  # add axis title and suffix to tick labels
                    yaxis = list(title = "diameter", ticksuffix = " m"), 
                    zaxis = list(title = "capacity", ticksuffix = " kW")), showlegend = FALSE)
```

**Click and drag on the plot below to move the axes.**  
<br>
```{r plotly 3d output, echo = FALSE}
shiny::div(turbine_3d)
```


<br>
<br>
<br>


## Location of turbines

Geographical data for each turbine is provided. This can be used to explore the location of wind farms.

```{r state data, results='asis'}
raw_data %>% 
  count(State = t_state) %>% 
  arrange(desc(n)) %>% 
  head(10) %>%
  xtable() %>% 
  print(type = "html", include.rownames = FALSE, 
        html.table.attributes = list('align="center", border=0'))
```


### Plot of locations

```{r map plot}
library(maps)

states = map_data("state")  # data for plotting US map

ggplot() + 
      geom_polygon(data = states, 
                   aes(x = long, y = lat, group=group), fill = NA, color = "black") + 
      geom_point(data = raw_data, 
                 aes(x = xlong, y = ylat), color = "blue3", size = 0.7) +
      coord_fixed(1.3) + xlim(range(states$long)) + ylim(range(states$lat)) + theme_void()

ggplot() + 
      geom_polygon(data = states, 
                   aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
      stat_density2d(data = raw_data, 
                     aes(x = xlong, y = ylat, fill = ..level..), geom = "polygon") +  # add density to map
      coord_fixed(1.3) + xlim(range(states$long)) + ylim(range(states$lat)) + theme_void()
```


### Animation {#Animation}

<a href="https://github.com/thomasp85/gganimate/wiki" target="_blank">gganimate</a>
is a package which can be used to create animations in R. It's quite simple to use; it only requires a few extra functions on top of a standard ggplot to turn it into an animation.  

```{r map animation, fig.align='center', fig.width=15}
library(gganimate)

animation_data <- raw_data %>% filter(!is.na(p_year))

location_animation <- ggplot() + 
      geom_polygon(data = states, 
                   aes(x = long, y = lat, group = group), 
                   fill = NA, color = "black") +  # add outline of US and states
      geom_point(data = animation_data, 
                 aes(x = xlong, y = ylat), 
                 color = "blue3", size = 0.7) +  # turbine location plot
  coord_fixed(1.3) + 
  xlim(range(states$long)) + 
  ylim(range(states$lat)) + 
  theme_void() + 
  labs(title = "Location of wind turbines in the USA",
       subtitle = 'Year: {frame_time}') +
  # animations settings, shdow_mark sticks points onto the map
  transition_time(p_year) + 
  shadow_mark(colour = 'blue')  

 # create animation
animate(location_animation, 
        nframes = length(unique(raw_data$p_year)), 
        fps = 1, 
        width = 800, height = 800)
```

#### Other animation

Having looked a gganimate, I spotted this tweet from
<a href="https://twitter.com/drob" target="_blank">@drob</a>
where he makes a very similar animation but with a few tweaks. It's interesting to see a similar but alternative approach.  

<blockquote class="twitter-tweet" data-cards="hidden" data-lang="en"><p lang="en" dir="ltr">Here&#39;s the graph I was aiming for in yesterday&#39;s <a href="https://twitter.com/hashtag/tidytuesday?src=hash&amp;ref_src=twsrc%5Etfw">#tidytuesday</a> screencast. This kind of data is a great use case for gganimate <a href="https://t.co/3qsndkMFZn">https://t.co/3qsndkMFZn</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://t.co/6LLfiEdUqs">pic.twitter.com/6LLfiEdUqs</a></p>&mdash; David Robinson (@drob) <a href="https://twitter.com/drob/status/1060191453874401280?ref_src=twsrc%5Etfw">November 7, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

```{r drob animation}
us_wind_processed <- raw_data %>%
  filter(!t_state %in% c("AK", "HI", "GU", "PR"))

wind_projects <- us_wind_processed %>%
  group_by(p_name, t_state) %>%
  summarize(year = min(p_year, na.rm = TRUE),
            turbines = n(),
            total_capacity = sum(t_cap),
            lon = mean(xlong),
            lat = mean(ylat),
            lon_sd = sd(xlong),
            lat_sd = sd(ylat)) %>%
  ungroup()

wind_projects %>%
  filter(!is.na(year), !is.infinite(year)) %>% 
  crossing(time = 1981:2018) %>%
  filter(year <= time) %>%
  ggplot(aes(lon, lat, size = turbines, color = year)) +
  borders("state") + 
  geom_point() +
  transition_manual(time) + 
  scale_color_continuous(guide = FALSE) + 
  labs(title = "Locations of wind turbine projects in continental US (1981-2018)") + 
  coord_map() +
  theme_void()
```
