---
title: "Week 30 - 2018"
---

***

## Movie Data
<br>

[Week 30](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-10-23) of tidy Tuesday is data from
<a href="https://www.the-numbers.com/" target="_blank">the-numbers.com</a>, it documents movies and their profit/cost.

<br>

### Plotly

The <a href="https://plot.ly/r/" target="_blank">plotly</a> library creates interactive graphs. Graphs can be created using base R graphics or using
<a href="https://ggplot2.tidyverse.org/" target="_blank">ggplot2</a>.  

The graph below shows the median production cost of movies by genre and year. Genre's can be hidden interactively.  
<br>


```{r load and tidy, warning=FALSE, message=FALSE, error=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(scales)
library(plotly)

 # Read data
horror_movie <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-10-23/movie_profit.csv")

 # Tidy dates and calculate return
horror_movie <- horror_movie %>% select(-X1) %>%
  mutate(release_date = lubridate::mdy(release_date), 
         release_year = lubridate::year(release_date),
         return = worldwide_gross - production_budget)

  # Aggregate and summarise data by year and genre
horror_movie %>% 
  group_by(release_year, genre) %>%
  summarise(avg_return = mean(return), 
            median_return = median(return),
            avg_production = mean(production_budget), 
            median_production = median(production_budget),
            avg_gross = mean(worldwide_gross), 
            median_gross = median(worldwide_gross)) -> movie_agg
```


```{r main plotly, fig.align='center', echo=FALSE}
  # ggplot of median production cost versus year by genre
yearly_cost_plot <- movie_agg %>% 
  ggplot(aes(release_year, median_production, colour=genre)) +
  geom_line() +
  scale_y_continuous(labels = dollar_format(scale=0.000001, suffix="M")) +  # Change y-axis labels to dollars
  ylab(NULL) + xlab("Year") + 
  labs(colour=NULL) +  # remove legend title
  ggtitle("Median Production Cost")

 # Create plotly graph
yearly_cost_plotly <- ggplotly(yearly_cost_plot) %>% 
  layout(margin = list(l = 45))  # add extra space on left for axis labels

shiny::div(yearly_cost_plotly, 
           align = "center")  # need the div function to centre the graph on the webpage
```

<br>

## Code

### Loading and tidying data

```{r ref.label="load and tidy", eval=FALSE}
```


### ggplot graph

```{r graph data, fig.align='center'}
  # ggplot of median production cost versus year by genre
yearly_cost_plot <- movie_agg %>% 
  ggplot(aes(release_year, median_production, colour=genre)) +
  geom_line() +
  scale_y_continuous(labels = dollar_format(scale=0.000001, suffix="M")) +  # Change y-axis labels to dollars
  labs(title = "Median Production Cost", x = "Year", y = NULL, colour = NULL)  # remove legend title

yearly_cost_plot
```


### plotly graph

After creating a ggplot graph, the `ggplotly` function will create a plotly version of the ggplot.

```{r plotly graph, fig.align='center'}
 # Create plotly graph
yearly_cost_plotly <- ggplotly(yearly_cost_plot) %>% 
  layout(margin = list(l = 45))  # add extra space on left for axis labels

shiny::div(yearly_cost_plotly, align = "center")  # need the div function to centre the graph on the webpage
```