---
title: "Week5"
---

***

The data for tidy tuesday week 5 is [US census data](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-04-30). The data contains information at a county level for a number of different variables/factors, such as population, income, and employment. More information on the dat acan be found [here](https://www.kaggle.com/muonneutrino/us-census-demographic-data).

<br>
<div align="center">
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">For <a href="https://twitter.com/hashtag/TidyTuesday?src=hash&amp;ref_src=twsrc%5Etfw">#TidyTuesday</a> I created simple function which allows you to plot any continuous variable in the data on a map <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://twitter.com/hashtag/r4ds?src=hash&amp;ref_src=twsrc%5Etfw">#r4ds</a> <a href="https://t.co/6Q1I121VqI">pic.twitter.com/6Q1I121VqI</a></p>&mdash; Aidan Boland (@AidoBo) <a href="https://twitter.com/AidoBo/status/991338257391804416?ref_src=twsrc%5Etfw">May 1, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
</div>
<br>

## Code

### Load libraries and data

```{r libraries and load data, warning=FALSE, message=FALSE, error=FALSE}
library(dplyr)
library(ggplot2)
library(maps)

tidy_tuesday5 <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-04-30/week5_acs2015_county_data.csv")
```


### County plot

I decided to create a function which would plot the data on a map. The function takes any of the numerical variables in the data and will output a heatmap of that variable.

```{r county data, fig.align='center'}
 # County map data
counties <- map_data("county")

 # Combined county data
all_county <- tidy_tuesday5 %>% 
  mutate(County=tolower(County), State=tolower(State)) %>%
  inner_join(counties, by=c("County" = "subregion", "State" = "region"))

county_plot <- function(x){
  ggplot(data = all_county, mapping = aes(x = long, y = lat, group = group)) + 
    coord_fixed(1.3) + 
    geom_polygon(aes_string(fill = x), color = "grey", size=0.05) +
    scale_fill_distiller(palette = "Spectral") +
    theme_void() 
}

county_plot("Income")
county_plot("MeanCommute")
```



### State version

I also created a version to plot data on a state level
```{r state function, fig.align='center'}
 # State map data
states <- map_data("state")

 # Aggregated state data
all_state <- tidy_tuesday5 %>%
  group_by(State = State) %>%
  summarise(Pop = sum(TotalPop),
            Income = mean(Income, na.rm = T),
            Men = sum(Men), Women = sum(Women),
            Unemployment_tot = sum(Unemployment * 0.01 * TotalPop)) %>%
  mutate(Ratio_men = Men / (Men + Women), 
         Unemployment = Unemployment_tot/Pop, 
         State = tolower(State)) %>%
  inner_join(states, by = c("State" = "region"))


 # Function to plot variable
state_plot <- function(x){
  ggplot(data = states, mapping = aes(x = long, y = lat, group = group)) +
    coord_fixed(1.3) +
    geom_polygon(data = all_state, aes_string(fill = x), color = "white") +
    geom_polygon(color = "black", fill = NA) +
    scale_fill_distiller(palette = "Spectral") +
    theme_void()
}


state_plot("Pop")
state_plot("Ratio_men")
state_plot("Income")
state_plot("Unemployment")
```



